\documentclass{article}
\usepackage[margin=.5in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{amsthm}
\usepackage{setspace}
\usepackage{esint}
\usepackage{amssymb}
\setstretch{.9}

\newenvironment{andrew_section}[1]
    {
    \section{#1}
    \begin{itemize}
    }
    {
    \end{itemize}
    }

\title{5 - Pseudosphere and the hyperbolic plane}

\begin{document}

\maketitle

\begin{andrew_section}{Belatrami's Insight}
    \item 
        two areas of discovery were reaching a head at the same time.
        hyperbolic geometry by lobachevsky, and gauss' differential geometry.
        both around 1830.
    \item 
        1868, belrimi thought there might be a connection. hyperbolic
        geometry was 40 years old and obscure.  [in some way, this makes
        sense to me- he was legit just denying one of the euclidean axioms
        and showing it produced cool stuff].
    \item 
        beltrimi reasoned via the local gauss-bonnet theorom
        that if he found a surface of constant negative curvature
        it would obey the laws of hyperbolic geometry.
\end{andrew_section}

\begin{andrew_section}{The Tractrix and the pseudosphere}
    \item 
        psuedosphere has this constant negative curvature, but it has
        a boundary...which is something of a problem because the hyperbolic
        plane was conceived to be infinite
    \item 
        \emph{tratrix}, investigated by newton.  imagine a table (axes X and Y),
        with a paper weight on it, and a string of length R attatched to it.
        the other end of the string touches the end of the table.  drag the
        string around the \emph{edge} of the table. curve the weight makes
        is a tractrix.  the weight will hypothetically never reach the 'Y-axis'.
    \item 
        let $\sigma$ represent the arc length, so $\sigma=0$ would be
        before you start pulling the string at all.  [nifty triangle similarty
        argument between delta sigma/-delta X and X/R...second triangle being
        the big one made by the string and distance from Y axis, first one
        obviously being infinitesimal]: \\ $X = Re^{-\frac{\sigma}{R}}$
    \item 
        pseudosphere is the surface of revolution for a tratrix
\end{andrew_section}

\begin{andrew_section}{A Conformal Map of the Pseudosphere}
    \item 
        need to create a conformal map from pseudosphere to $\mathbb{C}$.
        let x be the angle and sigma the arc length of the tractrix.
        so $(x,\sigma)$ will be sent to $x + yi$
    \item 
        a few considerations lock you into the bounded circle approach.
    \item 
        generators are orthogonal so the mapping must be orthogonal.
    \item 
        take a circle around the pseduosphere of sigma = constant.  this
        circle will have a radius X.  now take a triangle of two lines Xdx,
        $d\sigma$, because of the metric, this has to be shrunk to a trianle of
        lines dx dsigma.
    \item 
        the map is $d\hat{s} = \frac{R ds}{y}$, where $ds = \sqrt{dx^2 + dy^2}$
\end{andrew_section}

\begin{andrew_section}{the beltrami-poincare half-plane}
    \item 
        it is typical to set the radius of the pseudosphere
        to 1 so the curvature is -1.
    \item 
        it was important to beltrami that the pseduosphere be generalized
        to be infinite, like real hyperbolic geometry.  imagine a paint roller
        of radius 1.  after one revolution on a wall you have painted a strip
        $2\pi$ thick.  now we use our pseudosphere as a roller on the plane.
        horizontal strips represent lines of constant$\sigma$, going around
        the one circular strip of the pseduosphere.
    \item 
        that removes the lack of inifinitude, but it still has a rim that
        must be removed.  one can easily use the conformal mapping and the known
        metric to extend the rim further - you know exactly the scale factor
        thanks to the metric so you just keep going.
    \item 
        standardized hyperbolic metric: $d\hat{s} = \frac{ds}{y}$
    \item 
        this really makes for a half-plane.  the tractrix's 'end' at 
        y = 0.  if you imagine a particle traveling in this space,
        it will slow down more and more the closer you get to
        y = 0.
    \item 
        thus we now have \emph{The Hyperbolic Plane} $\mathbb{H}^2$,
        defined on y > 0, with the metric above.
    \item 
        to imagine how the metric affects things like arc length,
        we can imagine painting the half plane with circles of radius
        $\epsilon$, and letting them scale.  the length of the path
        is how many balls it goes through - way more balls towards the origin
    \item   
        this all makes sense - but how can we explain how geodesics
        are semicircles that hit the original at a right angle?
\end{andrew_section}

\begin{andrew_section}{using optics to find geodesics}
    \item 
        light takes the path that minimizes time.  imagine points a and b.
        light starts at a and ends at b.  a is above water, b is below.
        now imagine the point where the line will hit the water.
        changing the position will change the time spent on the path.
        because of fermat's principle, the light will in general 
        change angles when it his the water...aka snells law.
    \item 
        if you nudge the (stationary) point where light hits by $\epsilon$,
        you increase the 'time in flight' by 
        $ \frac{\epsilon \sin{\theta_1}}{v_1}$ (extra time spent in air), 
        but decrease it by $ \frac{\epsilon \sin{\theta_2}}{v_2}$. noting
        that at a stationary point these two quantities must be equal, and
        cancelling the $\epsilon$, we have : \\
        $\frac{ \sin{\theta_1}}{v_1} = \frac{ \sin{\theta_2}}{v_2}$, a 
        rearranged snells law.
    \item 
        take $\hat{a}$ and $\hat{b}$ on pseudosphere, corresponding to
        a and b on $\mathbb{H}^2$.  imagine a particle taking different
        paths between the points but always at unit speed. on hyperbolic
        plane, because of the metric, the particle will not travel at unit
        speed. bc of conformality, this change in speed doesn't depend on direction
    \item 
        on pseudosphere, geodesic will be the shortest path, and given the constant
        velocity...also shortest time, thus obeying fermat's principle
    \item 
        thus geodesics in the half plane follow snells law, which means
        they follow semicircle paths.
\end{andrew_section}

\begin{andrew_section}{the angle of parallelism}
    \item 
        now it's obvious, visually, that there are infinitly many
        parallel lines given a point p and a line L.  they are called
        \emph{ultra-parallel}
    \item 
        looking at same diagram, [5.9], for any point p and 
        line L, there are also 2 lines that only hit L at the horizon,
        called 'asymytotic'
    \item 
        given this point p and a line L, there is still only one point (q)
        that, when connected by a geodesic, will hit L at a right angle, we
        will call this line M. M lets us define a distance from a point to a line.
    \item 
        take this setup as a whole - for any point p and line L, we have 
        two asmptotic lines, and a third line M, hitting L at a right angle.
        M actually bisects the angle contained by the two asymtotics.
        angle between M and an asmyptotic is called the angle of parallelism,
        usually called $\Pi$.  
    \item 
        distance between p and L is denoted D.  there is a relationship
        between D and $\Pi$, $\tan({\frac{\Pi}{2}}) = e^{-D}$, 
        Bolyai-Lobachevski Formula.
    \item 
        as p approaches L, $\Pi$ approaches $\frac{\pi}{2}$, looks euclidean.
        as p gets farther away from L, the proportion of rays eminating from p
        that hit L become smaller and smaller.
    \item 
        to a tiny observer, no hyperbolic line is different from any other.
        to a global observer, it seems like some lines only have one endpoint
        and then go off to infinity.  in actuality, those lines all meet at
        infinity.  shows how important it is to switch between the different
        perspectives
    \item 
        meaningful application of this: rotating something about a point.
        distance preserving transformations are called isometries.  thinking
        about a transformation of an arc looking hyperbolic line, we can rotate
        it until it becomes a verticle line.
\end{andrew_section}

\begin{andrew_section}{the beltrami-poincare disc}
    \item 
        half plane is one way of representing $\mathbb{H}^2$, but there
        are other (mathematically but not psychologically) equivelent
        representations.  one is the beltrami-poincare disc.
    \item 
        infinite distant horizon is a boundary disk, metric is: \\
        $d\hat{s} = \frac{2}{1 - r^2} ds$, r is distance from center of disc
\end{andrew_section}

\end{document}